---
title: "Homework 6: number 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


2. Let’s consider the ARIMA(1,1,0) model with φ1 = 0.8. Use the R function arima to simulate a time series of length n = 500. Use the R set.seed(7) function. Make a time series plot of the data, sample ACF plot, and sample PACF. (See the example and Lab.) Make a time series plot of the differenced data, sample ACF plot, and sample PACF. (See the Lab.) From your plots of the data, what model would you fit to the data?


```{r fig.height=7, fig.width=8}

set.seed(7)

y <- arima.sim(model=list(order=c(1,1,0), ar=c(0.8)), n=500)

par(mfrow=c(3,2))

plot(y)
plot(diff(y))
acf(y)
acf(diff(y))
acf(y, type = "partial")
acf(diff(y), type = "partial")

```

From the plot of different series, PACF is cut off after lag 1 indicating an AR(1) model. The plot of different y seems like it's stationary. And sample ACF is exponential decay. I would fit it using an ARIMA(1,0,0)

