---
title: "Homework 8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


3) U.S. Unemployment Rate Data: The U.S. unemployment rate series can be found in the file in ur.dat off the class web page at https://edoras.sdsu.edu/~babailey/stat673/ur.dat. 

```{r}

# To read this file into an R dataset and make it an R time series object use the R commands:

ur <- scan("https://edoras.sdsu.edu/~babailey/stat673/ur.dat")
ur <- ts(ur, start=1948, frequency=4)

```


(a) Plot the time series, sample ACF and sample PACF.

```{r fig.height=3, fig.width=8}

par(mfrow=c(1,3))

plot(ur)
acf(ur)
acf(ur, type = "partial")

```

(b) Is an AR(2) model appropriate for the data? Explain. 

Since ACF is exponential decaying and PACF is cutoff after lag two, AR(2) can be an appropriate model the data. However, the time series is non-stationary. It seems like the unemployment rate is increasing over time.   


(c) Fit an AR(2) model to the data using the R function `arima`.

```{r echo = TRUE}

ur.fit1 <- arima(ur, order = c(2,0,0)) 

ur.fit1

```


(d) Make diagnostics plots using the R function tsdiag. How well does the model fit the data?

```{r fig.height=6, fig.width=5}

tsdiag(ur.fit1)

```

The standard residuals looks like there is some pattern with bigger residuals at some years and smaller at other years. ACF of residuals looks like white noise which is good. P-values plot looks decent at lag 2 to 6 then nearly 0 after lag 6 suggesting the model is not fitting quite well. 

(e) Use the first 115 data points to fit an AR(2) model. Forecast the next 6 observations and construct the 95% forecast limits. Include the plot. Compare your forecasts with the observations and quantify the difference.

```{r}

# split the training set 
ur1 <- ts(ur[1:115])

# fit the first 115 observation points 
ur.fit2 <- arima(ur1, order=c(2,0,0))

# and compute forecasts for the next 6 observation 
pred <- predict(ur.fit2, n.ahead = 6)

# graph the forecast with actual points 
plot(c(ur1, pred$pred), ylim=c(0,13), xlim = c(0, 130), type="n", ylab = "U.S. Unemployment Rate")
lines(1:115, ur1, lty=1)
lines(116:121, pred$pred + 2*pred$se, lty=2, col = "red")
lines(116:121, pred$pred - 2*pred$se, lty=2, col = "red")
points(116:121, pred$pred, pch = 3, col = "blue")
points(116:121, ur[116:121], pch=1, col = "green") 
title("95% Forecast intervals")

legend("topleft", c("Forecases", "Actual UR", "95% Forecast CI"), 
       pch = c(3, 1, NA), lty = c(NA, NA, 2), col = c("blue", "green", "red"))

```

The forecasts and actual unemployment rate for these 6 observations are very close. This means that the model is predicting these test points pretty well. Since the time series is non-stationary, the 95% confident interval will be wider as we're making more predictions.  


